Compositional Certified Resource Bounds\ifshort{\vspace{-.15cm}}{}

### 2014-11-11 paper/main.tex
section shuffling
integration of the example derivation

\begin{figure*}[t!]
 \setlength{\progwidth}{.22\linewidth}
  \centering
\hspace{-0.4cm}
  \begin{minipage}[b]{.18\linewidth}
    \begin{center}
   \begin{lstlisting}[]
while (n>x) {
  ${\color{blue}%
  n{>}x; \, |[x, n]| {+} |[y, m]| \vdash }$
  if (m>y)
    ${\color{blue}%
    m{>}y; \, |[x, n]| {+} |[y, m]| \vdash }$
    y=y+1;
    ${\color{blue}%
    \dashv \cdot ; \, 1 {+} |[x, n]| {+} |[y, m]| }$
  else
    ${\color{blue}%
    n{>}x; \, |[x, n]| {+} |[y, m]| \vdash }$
    x=x+1;
    ${\color{blue}%
    \dashv \cdot; \, 1 {+} |[x, n]| {+} |[y, m]| }$
  ${\color{blue}%
  \dashv \cdot; \, 1 {+} |[x, n]| {+} |[y, m]| \vdash }$
  tick(1);
  ${\color{blue}%
  \dashv \cdot; \, |[x, n]| {+} |[y, m]| }$
}
   \end{lstlisting}

@@

\begin{figure*}[t]

@@

\sectskip
\section{Automatic Inference via LP Solving.}
\label{sec:lp}
\aftersectskip

We separate the search of a derivation into two steps.  As a first
step we go through the whole program and apply inductively the
derivation rules of the automatic amortized analysis.  During this
process our tool uses symbolic names for the rational coefficients
$(q_i)$ in the rules.  Every time a linear constraint must be
satisfied by these coefficients, it is recorded in a global list using
the symbolic names.  We then feed the collected constraints to an
off the shelf LP solver~\footnote{We currently use Coin-Or's CLP.}.
If the solver successfully finds a solution, we know that a derivation
exists for the considered program.  We can then extract the initial
$Q$ from the solver and get a resource bound for the program.  To get
a full derivation we simply extract the complete solution from the
solver, and apply it to the symbolic names $(q_i)$ of the coefficients
in the derivation.  If the LP solver fails to find a solution, an
error is reported to the user.

@@

\ifshort{
\begin{figure*}[t]
\small
\begin{mathpar}
%
\Rule{L:Skip}
{ }
{ \Delta; B; R \vdash
\htriple
  { Q }{ \code{skip} }{ Q }
}
%
\and
\Rule{L:Break}
{ }
{ \Delta; B; R \vdash
\htriple
  { M_b + B }{ \code{break} }{ Q }
}
%
\and
\Rule{L:Return}
{ }
{ \Delta; B; R \vdash
\htriple
  { R\,(\state(x)) }{ \code{return}~x }{ Q }
}
%
\and
\Rule[leftskip=.0cm,rightskip=.0cm]{L:Update}
{ }
{ \Delta; B; R \vdash
\htriple
  { \lambda \state.\,M_u + M_e(e) + Q\,\state[x \mapsto \sem{e}_\state] }{ x \gets e }{ Q }
}
%
\and
\Rule[leftskip=.0cm,rightskip=.0cm]{L:Seq}
{ \Delta; B; R \vdash
\htriple{ P }{ S_1 }{ Q' + M_s }
\\
\!\!  \Delta; B; R \vdash
\htriple{ Q' }{ S_2 }{ Q }
}
{ \Delta; B; R \vdash
\htriple{ P }{ S_1;S_2 }{ Q }
}
%
\and
\Rule{L:Call}
{
  \Delta(f) = \forall z\,\vec v\,v.(P_f\,z\,\vec v, Q_f\,z\,v) \\
  P \models P_f \, y \, (\state(\vec x)) \land A \\
  \forall v.\, (Q_f \, y \, v \land A \models \lambda \state.\, Q \, \state[r \mapsto v])
}
{ \Delta; B; R \vdash
\htriple
  { M_f + P }
  { r \gets f(\vec x) }
  { Q - M_r }
}
%
\and
\Rule{L:Assert}
{ }
{ \Delta; B; R \vdash
\htriple
  { \code{istrue}~\sem{e}_\state \implies Q + M_a }
  { \code{assert}~e }
  { Q }
}
%
\and
\Rule{L:Tick}
{ }
{ \Delta; B; R \vdash
\htriple
  { Q + M_t(n) }
  { \code{tick}(n) }
  { Q }
}
%
\and
\Rule{L:Loop}
{ \Delta; Q; R \vdash
\htriple
  { I }{ S }{ I + M_l }
}
{ \Delta; B; R \vdash
\htriple
  { I }{ \code{loop}~S }{ Q }
}
%
\and
\Rule{L:If}
{ \Delta; B; R \vdash
\htriple
  { \code{istrue}~\sem{e}_\state + P - M_c^1}
  { S_1 }{ Q }
\\
  \Delta; B; R \vdash
\htriple
  { \code{isfalse}~\sem{e}_\state + P - M_c^2 }
  { S_2 }{ Q }
}
{ \Delta; B; R \vdash
\htriple
  { P + M_e(e) }{ \code{if}(e)~S_1~\code{else}~S_2 }{ Q }
}
%
\and
\Rule{L:Weaken}
{
  \!\! P \models P' \!\! \\
  \!\! \Delta; B'; R' \vdash
  \htriple{ P' }{ S }{ Q' } \!\! \\
  \!\! Q' \models Q \!\! \\
  \!\! B' \models B \!\! \\
  \!\! \forall v.\, (R'\,v \models R\,v) \!\!
}
{ \Delta; B; R \vdash
\htriple{ P }{ S }{ Q }
}
%
\and
\Rule{L:Frame}
{ \Delta; B; R \vdash
\htriple{ P }{ S }{ Q } \\
  x \in \Qplusz
}
{ \!\! \Delta; B + x; R + x \vdash
\htriple{ P + x }{ S }{ Q + x } \!\!
}
%
\and
\Rule{L:Extend}
{ \Delta \cup \Delta'; B; R \vdash
  \htriple{ P }{ S }{ Q }
\\
  \forall f\,P_f\,Q_f.\,
  \Delta'(f) = \forall z\,\vec v\,v.(P_f\,z\,\vec v, Q_f\,z\,v)
  \rightarrow
  \forall y\,\vec v.\,
  (\Delta \cup \Delta'; \bot; Q_f\,y \vdash
    \htriple{ P_f\,y\,\vec v }{ S_f }{ \bot })
}
{ \Delta; B; R \vdash
\htriple{ P }{ S }{ Q }
}
%
\end{mathpar}

@@

\iffull{
\begin{figure*}[t]
\begin{mathpar}
%
\Rule{L:Skip}
{ }
{ \Delta; B; R \vdash
\htriple
  { Q }{ \code{skip} }{ Q }
}
%
\and
\Rule{L:Break}
{ }
{ \Delta; B; R \vdash
\htriple
  { M_b + B }{ \code{break} }{ Q }
}
%
\and
\Rule{L:Return}
{ }
{ \Delta; B; R \vdash
\htriple
  { R\,(\state(x)) }{ \code{return}~x }{ Q }
}
%
\and
\Rule[leftskip=.0cm,rightskip=.0cm]{L:Update}
{ }
{ \Delta; B; R \vdash
\htriple
  { \lambda \state.\,M_u + M_e(e) + Q\,\state[x \mapsto \sem{e}_\state] }{ x \gets e }{ Q }
}
%
\and
\Rule[leftskip=.0cm,rightskip=.0cm]{L:Seq}
{ \Delta; B; R \vdash
\htriple{ P }{ S_1 }{ Q' + M_s }
\\
\!\!  \Delta; B; R \vdash
\htriple{ Q' }{ S_2 }{ Q }
}
{ \Delta; B; R \vdash
\htriple{ P }{ S_1;S_2 }{ Q }
}
%
\and
\Rule{L:Call}
{
  \Delta(f) = \forall z\,\vec v\,v.(P_f\,z\,\vec v, Q_f\,z\,v) \\
  P \models P_f \, y \, (\state(\vec x)) \land A \\
  \forall v.\, (Q_f \, y \, v \land A \models \lambda \state.\, Q \, \state[r \mapsto v])
}
{ \Delta; B; R \vdash
\htriple
  { M_f + P }
  { r \gets f(\vec x) }
  { Q - M_r }
}
%
\and
\Rule{L:Assert}
{ }
{ \Delta; B; R \vdash
\htriple
  { \code{istrue}~\sem{e}_\state \implies Q + M_a }
  { \code{assert}~e }
  { Q }
}
%
\and
\Rule{L:Tick}
{ }
{ \Delta; B; R \vdash
\htriple
  { Q + M_t(n) }
  { \code{tick}(n) }
  { Q }
}
%
\and
\Rule{L:Loop}
{ \Delta; Q; R \vdash
\htriple
  { I }{ S }{ I + M_l }
}
{ \Delta; B; R \vdash
\htriple
  { I }{ \code{loop}~S }{ Q }
}
%
\and
\Rule{L:If}
{ \Delta; B; R \vdash
\htriple
  { \code{istrue}~\sem{e}_\state + P - M_c^1}
  { S_1 }{ Q }
\\
  \Delta; B; R \vdash
\htriple
  { \code{isfalse}~\sem{e}_\state + P - M_c^2 }
  { S_2 }{ Q }
}
{ \Delta; B; R \vdash
\htriple
  { P + M_e(e) }{ \code{if}(e)~S_1~\code{else}~S_2 }{ Q }
}
%
\and
\Rule{L:Weaken}
{
  \!\! P \models P' \!\! \\
  \!\! \Delta; B'; R' \vdash
  \htriple{ P' }{ S }{ Q' } \!\! \\
  \!\! Q' \models Q \!\! \\
  \!\! B' \models B \!\! \\
  \!\! \forall v.\, (R'\,v \models R\,v) \!\!
}
{ \Delta; B; R \vdash
\htriple{ P }{ S }{ Q }
}
%
\and
\Rule{L:Frame}
{ \Delta; B; R \vdash
\htriple{ P }{ S }{ Q } \\
  x \in \Qplusz
}
{ \!\! \Delta; B + x; R + x \vdash
\htriple{ P + x }{ S }{ Q + x } \!\!
}
%
\and
\Rule{L:Extend}
{ \Delta \cup \Delta'; B; R \vdash
  \htriple{ P }{ S }{ Q }
\\
  \forall f\,P_f\,Q_f.\,
  \Delta'(f) = \forall z\,\vec v\,v.(P_f\,z\,\vec v, Q_f\,z\,v)
  \rightarrow
  \forall y\,\vec v.\,
  (\Delta \cup \Delta'; \bot; Q_f\,y \vdash
    \htriple{ P_f\,y\,\vec v }{ S_f }{ \bot })
}
{ \Delta; B; R \vdash
\htriple{ P }{ S }{ Q }
}
%
\end{mathpar}

@@

\begin{figure*}[t]
  \centering
  \footnotesize
  \begin{mathpar}
  \inferrule*[right={\scriptsize (Q:Loop)},leftskip=0cm,rightskip=0cm,width=20cm]
  { \inferrule*[right={\scriptsize (Q:If)},leftskip=0cm,rightskip=0cm,width=20cm]
    {
      \inferrule*[right={\scriptsize (Q:Weak)},leftskip=2cm,rightskip=4cm, vdots=1.5cm]
      {
        \inferrule*[right={\scriptsize (Q:Seq)},leftskip=0cm,rightskip=0cm]
        {
\inferrule*[right={\scriptsize (Q:Weak)},leftskip=0cm,rightskip=0cm]
{
  \inferrule*[right={\scriptsize (Q:DecP)},leftskip=0cm,rightskip=0cm]
  { }
  {(x {<} 10, B^\text{de}); ((x {\geq} 10,Q^\text{de})  \vdash \code{x = x - 10}
   \dashv (\cdot, P^\text{de})
  }
}
{(x {<} 10, B^\text{we}); ((x {\geq} 10,Q^\text{we})  \vdash \code{x = x - 10}
\dashv (\cdot, P^\text{we})
}
\and
\inferrule*[right={\scriptsize (Q:Tick)},leftskip=0cm,rightskip=0cm]
{ }
{(x {<} 10, B^\text{ti}); (\cdot,Q^\text{ti})  \vdash \code{tick(5)}
 \dashv (\cdot, P^\text{ti})
}
} {
          (x {<} 10, B^\text{sq}); ((x {\geq} 10,Q^\text{sq}) \vdash \code{x=x{-}10; tick(5);} \dashv (\cdot,P^\text{sq})
        }
      }
      { (x {<} 10, B^\text{if}); ((x {\geq} 10,Q^\text{if}) \vdash \code{x=x{-}10; tick(5);} \dashv (\cdot,P^\text{if}) }
      \and
      \inferrule*[right={\scriptsize (Q:Weak)},leftskip=4.5cm,rightskip=4cm]
      { \inferrule*[right={\scriptsize (Q:Break)},leftskip=0cm,rightskip=0cm]
        { }
        {(x {<} 10, B^\text{br}); (x {<} 10,Q^\text{br}) \vdash \code{break} \dashv (\bot,P^\text{br})}
      }
      {
        (x {<} 10, B^\text{el}); (x {<} 10,Q^\text{el}) \vdash \code{break} \dashv (\cdot,P^\text{el})
      }
    }
    { \hspace{3cm} (x {<} 10, B^\text{lo}); (\cdot,Q^\text{lo}) \vdash \loopbody \dashv (\cdot,P^\text{lo}) \hspace{3cm}
    }
  }
  {
   (\cdot, B); (\cdot, Q) \vdash \code{loop}~{\loopbody} \dashv (x {<} 10, P)
  }
\end{mathpar}
\flushleft Constraints:
$$
\begin{array}{llll}
P = B^\text{lo} & Q  = Q^\text{lo} & Q^\text{lo} = P^\text{lo} \\
B^\text{el} = B^\text{if} = B^\text{lo} & Q^\text{el} = Q^\text{if} = Q^\text{lo} & P^\text{el} = P^\text{if} = P^\text{lo} \\
B^\text{el} = B^\text{br}  & Q^\text{el} \succeq_{(x{<}10)} Q^\text{br}  & P^\text{br} \succeq_{(\cdot)} P^\text{el} \\
B^\text{br} = Q^\text{br}\\
B^\text{if} = B^\text{sq}  & Q^\text{if} \succeq_{(x{<}10)} Q^\text{sq}  & P^\text{sq} \succeq_{(\cdot)} P^\text{if} \\
B^\text{sq} = B^\text{we} = B^\text{ti} & Q^\text{sq} = Q^\text{we} & P^\text{we} = Q^\text{ti} & P^\text{ti} = P^\text{sq} \\
Q^\text{ti} = P^\text{ti} + 5 \\
B^\text{we} = B^\text{de}  & Q^\text{we} \succeq_{(x{<}10)} Q^\text{de}  & P^\text{de} \succeq_{(\cdot)} P^\text{we} \\
p^\text{de}_{0,10} = q^\text{de}_{0,10} + q^\text{de}_{0,x}  & p^\text{de}_{0} = q^\text{de}_{0}
& \forall (\alpha, \beta) \neq (0,10)  . \, p^\text{de}_{\alpha,\beta} = q^\text{de}_{\alpha,\beta}
\end{array}
$$
\vspace{1.5ex}

\caption{An example derivation as produced by the automatic tool.}
\label{fig:derivation}
\end{figure*}
\sectskip
\section{Conclusion}
\label{sec:concl}
\aftersectskip

@@

\section{Example Derivation in the Automatic System}
\label{app:derivation}

@@

\begin{figure*}[t!]
\setlength{\progwidth}{.24\linewidth}
  \centering

@@

\begin{figure*}[t!]
 \setlength{\progwidth}{.24\linewidth}
  \centering
  \begin{minipage}[b]{\progwidth}
    \begin{center}
   \begin{lstlisting}
  while (n>x) {
    if (m>y)
      y = y+1;
    else
      x = x+1;
  }
   \end{lstlisting}

@@

\begin{figure*}[t!]
 \setlength{\progwidth}{.24\linewidth}
  \centering
  \begin{minipage}[b]{\progwidth}
    \begin{center}
   \begin{lstlisting}
  x=0;
  while (x<n) {
    if (*) break;
    x=x+1;
  }
  while (x<n)
    x=x+1;
   \end{lstlisting}

@@

\begin{figure*}[t!]
 \setlength{\progwidth}{.24\linewidth}
  \centering
%
%
  \begin{minipage}[b]{\progwidth}
    \begin{center}
   \begin{lstlisting}
  assert n>0;
  assert m>0;
  va = n; vb = 0;
  while (va>0 && *) {
    if (vb<m) {
      vb=vb+1;
      va=va-1;
    } else {
      vb=vb-1;
      vb=0;
    }
  }
   \end{lstlisting}
$1 + 2|[0, n]|$
\\[.7\baselineskip]
      {\bf fig4\_2}
    \end{center}
  \end{minipage}
%
%
%
%
  \begin{minipage}[b]{\progwidth}
    \begin{center}
   \begin{lstlisting}
  assert (0<m);
  i = n;
  while (i>0 && *) {
    if (i<m)
      i=i-1;
    else
      i=i-m;
  }
   \end{lstlisting}
$|[0, n]|$
\\[.7\baselineskip]
      {\bf fig4\_4}
    \end{center}
  \end{minipage}
%
%
  \begin{minipage}[b]{\progwidth}
    \begin{center}
   \begin{lstlisting}
  assert (0 < m < n);
  i=m;
  while (0<i<n) {
    if (dir==fwd) i++;
    else i--;
  }
   \end{lstlisting}
$---$
\\[.7\baselineskip]
      {\bf fig4\_5}
    \end{center}
  \end{minipage}

@@

\begin{figure*}[t!]
 \setlength{\progwidth}{.28\linewidth}
  \centering
%
%
  \begin{minipage}[b]{\progwidth}
    \begin{center}
   \begin{lstlisting}
void count_down (int x) {
  int a = x;
  if (a>0) {
    a = a-1;
    count_down(a);
  }
}
int copy (int x, int y) {
  if (x>0) {
    x = x-1;
    y = y+1;
    y=copy(x,y);
  };
  return y;
}
void main (int x,int y) {
  y = copy (x,y);
  count_down(y);
}
   \end{lstlisting}

>>>>

\begin{figure*}
 \setlength{\progwidth}{.22\linewidth}
  \centering
\hspace{-0.4cm}
  \begin{minipage}[b]{.18\linewidth}
    \begin{center}
   \begin{lstlisting}[]
while (n>x) {
  ${\color{blue}%
  n{>}x; \, |[x, n]| {+} |[y, m]| \vdash }$
  if (m>y)
    ${\color{blue}%
    m{>}y; \, |[x, n]| {+} |[y, m]| \vdash }$
    y=y+1;
    ${\color{blue}%
    \dashv \cdot ; \, 1 {+} |[x, n]| {+} |[y, m]| }$
  else
    ${\color{blue}%
    n{>}x; \, |[x, n]| {+} |[y, m]| \vdash }$
    x=x+1;
    ${\color{blue}%
    \dashv \cdot; \, 1 {+} |[x, n]| {+} |[y, m]| }$
  ${\color{blue}%
  \dashv \cdot; \, 1 {+} |[x, n]| {+} |[y, m]| \vdash }$
  tick(1);
  ${\color{blue}%
  \dashv \cdot; \, |[x, n]| {+} |[y, m]| }$
}
   \end{lstlisting}

@@

\begin{figure*}

@@

\sectskip
\section{Automatic Inference via LP Solving}
\label{sec:lp}
\aftersectskip

We separate the search of a derivation into two steps.  As a first
step we go through the whole program and apply inductively the
derivation rules of the automatic amortized analysis.  During this
process our tool uses symbolic names for the rational coefficients
$(q_i)$ in the rules.  Every time a linear constraint must be
satisfied by these coefficients, it is recorded in a global list using
the symbolic names.  We then feed the collected constraints to an
off the shelf LP solver. (We currently use Coin-Or's CLP.)
If the solver successfully finds a solution, we know that a derivation
exists for the considered program.  We can then extract the initial
$Q$ from the solver and get a resource bound for the program.  To get
a full derivation we simply extract the complete solution from the
solver, and apply it to the symbolic names $(q_i)$ of the coefficients
in the derivation.  If the LP solver fails to find a solution, an
error is reported to the user.

\begin{figure*}
\centering
\footnotesize
\begin{mathpar}
\inferrule*[right={\scriptsize (Q:Loop)},leftskip=0cm,rightskip=0cm,width=20cm]
{ \inferrule*[right={\scriptsize (Q:If)},leftskip=0cm,rightskip=0cm,width=20cm]
  { \inferrule*[right={\scriptsize (Q:Weak)},leftskip=2cm,rightskip=4cm, vdots=1cm]
    { \inferrule*[right={\scriptsize (Q:Seq)},leftskip=0cm,rightskip=0cm]
      { \inferrule*[right={\scriptsize (Q:Weak)},leftskip=0cm,rightskip=0cm]
        { \inferrule*[right={\scriptsize (Q:DecP)},leftskip=0cm,rightskip=0cm]
          {
          }
          { (x {<} 10, B^\text{de}); ((x {\geq} 10,Q^\text{de})  \vdash \code{x = x - 10}
            \dashv (\cdot, P^\text{de})
          }
        }
        { (x {<} 10, B^\text{we}); ((x {\geq} 10,Q^\text{we})  \vdash \code{x = x - 10}
          \dashv (\cdot, P^\text{we})
        }
        \and
        \inferrule*[right={\scriptsize (Q:Tick)},leftskip=0cm,rightskip=0cm]
        {
        }
        { (x {<} 10, B^\text{ti}); (\cdot,Q^\text{ti})  \vdash \code{tick(5)}
          \dashv (\cdot, P^\text{ti})
        }
      }
      { (x {<} 10, B^\text{sq}); ((x {\geq} 10,Q^\text{sq}) \vdash
        \code{x=x{-}10; tick(5);} \dashv (\cdot,P^\text{sq})
      }
    }
    { (x {<} 10, B^\text{if}); ((x {\geq} 10,Q^\text{if}) \vdash
      \code{x=x{-}10; tick(5);} \dashv (\cdot,P^\text{if})
    }
    \and
    \inferrule*[right={\scriptsize (Q:Weak)},leftskip=4.5cm,rightskip=4cm]
    { \inferrule*[right={\scriptsize (Q:Break)},leftskip=0cm,rightskip=0cm]
      { }
      { (x {<} 10, B^\text{br}); (x {<} 10,Q^\text{br}) \vdash
        \code{break} \dashv (\bot,P^\text{br})
      }
    }
    { (x {<} 10, B^\text{el}); (x {<} 10,Q^\text{el}) \vdash
      \code{break} \dashv (\cdot,P^\text{el})
    }
  }
  { \hspace{3cm} (x {<} 10, B^\text{lo}); (\cdot,Q^\text{lo}) \vdash
    \loopbody \dashv (\cdot,P^\text{lo}) \hspace{3cm}
  }
}
{ (\cdot, B); (\cdot, Q) \vdash \code{loop}~{\loopbody} \dashv (x {<} 10, P)
}
\end{mathpar}
\flushleft Constraints:
$$
\begin{array}{lll}
P {=} B^\text{lo} \land Q {=} Q^\text{lo} {=} P^\text{lo}
& B^\text{el} {=} B^\text{if} {=} B^\text{lo}
    \land Q^\text{el} {=} Q^\text{if} {=} Q^\text{lo}
    \land P^\text{el} {=} P^\text{if} {=} P^\text{lo}
& B^\text{el} {=} B^\text{br} \land Q^\text{el} \succeq_{(x{<}10)} Q^\text{br}
    \land P^\text{br} \succeq_{(\cdot)} P^\text{el}
\\
B^\text{br} {=} Q^\text{br}
& B^\text{if} {=} B^\text{sq} \land Q^\text{if} \succeq_{(x{<}10)} Q^\text{sq}
    \land P^\text{sq} \succeq_{(\cdot)} P^\text{if}
& B^\text{sq} {=} B^\text{we} {=} B^\text{ti} \land Q^\text{sq} {=} Q^\text{we}
    \land P^\text{we} {=} Q^\text{ti} \land P^\text{ti} {=} P^\text{sq}
\\
Q^\text{ti} {=} P^\text{ti} + 5
& B^\text{we} {=} B^\text{de} \land Q^\text{we} \succeq_{(x{<}10)} Q^\text{de}
    \land P^\text{de} \succeq_{(\cdot)} P^\text{we}
& p^\text{de}_{0,10} {=} q^\text{de}_{0,10} + q^\text{de}_{0,x}
    \land p^\text{de}_{0} {=} q^\text{de}_{0}
    \land \forall (\alpha, \beta) \neq (0,10)  . \,
              p^\text{de}_{\alpha,\beta} {=} q^\text{de}_{\alpha,\beta}
\end{array}
$$
\vspace{1.5ex}

\caption{An example derivation as produced by the automatic tool.}
\label{fig:derivation}
\end{figure*}

@@

\ifshort{
\begin{figure*}
\small
\begin{mathpar}
%
\Rule{L:Skip}
{ }
{ \Delta; B; R \vdash
\htriple
  { Q }{ \code{skip} }{ Q }
}
%
\and
\Rule{L:Break}
{ }
{ \Delta; B; R \vdash
\htriple
  { M_b + B }{ \code{break} }{ Q }
}
%
\and
\Rule{L:Return}
{ }
{ \Delta; B; R \vdash
\htriple
  { R\,(\state(x)) }{ \code{return}~x }{ Q }
}
%
\and
\Rule[leftskip=.0cm,rightskip=.0cm]{L:Update}
{ }
{ \Delta; B; R \vdash
\htriple
  { \lambda \state.\,M_u + M_e(e) + Q\,\state[x \mapsto \sem{e}_\state] }{ x \gets e }{ Q }
}
%
\and
\Rule[leftskip=.0cm,rightskip=.0cm]{L:Seq}
{ \Delta; B; R \vdash
\htriple{ P }{ S_1 }{ Q' + M_s }
\\
\!\!  \Delta; B; R \vdash
\htriple{ Q' }{ S_2 }{ Q }
}
{ \Delta; B; R \vdash
\htriple{ P }{ S_1;S_2 }{ Q }
}
%
\and
\Rule{L:Call}
{
  \Delta(f) = \forall z\,\vec v\,v.(P_f\,z\,\vec v, Q_f\,z\,v) \\
  P \models P_f \, y \, (\state(\vec x)) \land A \\
  \forall v.\, (Q_f \, y \, v \land A \models \lambda \state.\, Q \, \state[r \mapsto v])
}
{ \Delta; B; R \vdash
\htriple
  { M_f + P }
  { r \gets f(\vec x) }
  { Q - M_r }
}
%
\and
\Rule{L:Assert}
{ }
{ \Delta; B; R \vdash
\htriple
  { \code{istrue}~\sem{e}_\state \implies Q + M_a }
  { \code{assert}~e }
  { Q }
}
%
\and
\Rule{L:Tick}
{ }
{ \Delta; B; R \vdash
\htriple
  { Q + M_t(n) }
  { \code{tick}(n) }
  { Q }
}
%
\and
\Rule{L:Loop}
{ \Delta; Q; R \vdash
\htriple
  { I }{ S }{ I + M_l }
}
{ \Delta; B; R \vdash
\htriple
  { I }{ \code{loop}~S }{ Q }
}
%
\and
\Rule{L:If}
{ \Delta; B; R \vdash
\htriple
  { \code{istrue}~\sem{e}_\state + P - M_c^1}
  { S_1 }{ Q }
\\
  \Delta; B; R \vdash
\htriple
  { \code{isfalse}~\sem{e}_\state + P - M_c^2 }
  { S_2 }{ Q }
}
{ \Delta; B; R \vdash
\htriple
  { P + M_e(e) }{ \code{if}(e)~S_1~\code{else}~S_2 }{ Q }
}
%
\and
\Rule{L:Weaken}
{
  \!\! P \models P' \!\! \\
  \!\! \Delta; B'; R' \vdash
  \htriple{ P' }{ S }{ Q' } \!\! \\
  \!\! Q' \models Q \!\! \\
  \!\! B' \models B \!\! \\
  \!\! \forall v.\, (R'\,v \models R\,v) \!\!
}
{ \Delta; B; R \vdash
\htriple{ P }{ S }{ Q }
}
%
\and
\Rule{L:Frame}
{ \Delta; B; R \vdash
\htriple{ P }{ S }{ Q } \\
  x \in \Qplusz
}
{ \!\! \Delta; B + x; R + x \vdash
\htriple{ P + x }{ S }{ Q + x } \!\!
}
%
\and
\Rule{L:Extend}
{ \Delta \cup \Delta'; B; R \vdash
  \htriple{ P }{ S }{ Q }
\\
  \forall f\,P_f\,Q_f.\,
  \Delta'(f) = \forall z\,\vec v\,v.(P_f\,z\,\vec v, Q_f\,z\,v)
  \rightarrow
  \forall y\,\vec v.\,
  (\Delta \cup \Delta'; \bot; Q_f\,y \vdash
    \htriple{ P_f\,y\,\vec v }{ S_f }{ \bot })
}
{ \Delta; B; R \vdash
\htriple{ P }{ S }{ Q }
}
%
\end{mathpar}

@@

\iffull{
\begin{figure*}
\begin{mathpar}
%
\Rule{L:Skip}
{ }
{ \Delta; B; R \vdash
\htriple
  { Q }{ \code{skip} }{ Q }
}
%
\and
\Rule{L:Break}
{ }
{ \Delta; B; R \vdash
\htriple
  { M_b + B }{ \code{break} }{ Q }
}
%
\and
\Rule{L:Return}
{ }
{ \Delta; B; R \vdash
\htriple
  { R\,(\state(x)) }{ \code{return}~x }{ Q }
}
%
\and
\Rule[leftskip=.0cm,rightskip=.0cm]{L:Update}
{ }
{ \Delta; B; R \vdash
\htriple
  { \lambda \state.\,M_u + M_e(e) + Q\,\state[x \mapsto \sem{e}_\state] }{ x \gets e }{ Q }
}
%
\and
\Rule[leftskip=.0cm,rightskip=.0cm]{L:Seq}
{ \Delta; B; R \vdash
\htriple{ P }{ S_1 }{ Q' + M_s }
\\
\!\!  \Delta; B; R \vdash
\htriple{ Q' }{ S_2 }{ Q }
}
{ \Delta; B; R \vdash
\htriple{ P }{ S_1;S_2 }{ Q }
}
%
\and
\Rule{L:Call}
{
  \Delta(f) = \forall z\,\vec v\,v.(P_f\,z\,\vec v, Q_f\,z\,v) \\
  P \models P_f \, y \, (\state(\vec x)) \land A \\
  \forall v.\, (Q_f \, y \, v \land A \models \lambda \state.\, Q \, \state[r \mapsto v])
}
{ \Delta; B; R \vdash
\htriple
  { M_f + P }
  { r \gets f(\vec x) }
  { Q - M_r }
}
%
\and
\Rule{L:Assert}
{ }
{ \Delta; B; R \vdash
\htriple
  { \code{istrue}~\sem{e}_\state \implies Q + M_a }
  { \code{assert}~e }
  { Q }
}
%
\and
\Rule{L:Tick}
{ }
{ \Delta; B; R \vdash
\htriple
  { Q + M_t(n) }
  { \code{tick}(n) }
  { Q }
}
%
\and
\Rule{L:Loop}
{ \Delta; Q; R \vdash
\htriple
  { I }{ S }{ I + M_l }
}
{ \Delta; B; R \vdash
\htriple
  { I }{ \code{loop}~S }{ Q }
}
%
\and
\Rule{L:If}
{ \Delta; B; R \vdash
\htriple
  { \code{istrue}~\sem{e}_\state + P - M_c^1}
  { S_1 }{ Q }
\\
  \Delta; B; R \vdash
\htriple
  { \code{isfalse}~\sem{e}_\state + P - M_c^2 }
  { S_2 }{ Q }
}
{ \Delta; B; R \vdash
\htriple
  { P + M_e(e) }{ \code{if}(e)~S_1~\code{else}~S_2 }{ Q }
}
%
\and
\Rule{L:Weaken}
{
  \!\! P \models P' \!\! \\
  \!\! \Delta; B'; R' \vdash
  \htriple{ P' }{ S }{ Q' } \!\! \\
  \!\! Q' \models Q \!\! \\
  \!\! B' \models B \!\! \\
  \!\! \forall v.\, (R'\,v \models R\,v) \!\!
}
{ \Delta; B; R \vdash
\htriple{ P }{ S }{ Q }
}
%
\and
\Rule{L:Frame}
{ \Delta; B; R \vdash
\htriple{ P }{ S }{ Q } \\
  x \in \Qplusz
}
{ \!\! \Delta; B + x; R + x \vdash
\htriple{ P + x }{ S }{ Q + x } \!\!
}
%
\and
\Rule{L:Extend}
{ \Delta \cup \Delta'; B; R \vdash
  \htriple{ P }{ S }{ Q }
\\
  \forall f\,P_f\,Q_f.\,
  \Delta'(f) = \forall z\,\vec v\,v.(P_f\,z\,\vec v, Q_f\,z\,v)
  \rightarrow
  \forall y\,\vec v.\,
  (\Delta \cup \Delta'; \bot; Q_f\,y \vdash
    \htriple{ P_f\,y\,\vec v }{ S_f }{ \bot })
}
{ \Delta; B; R \vdash
\htriple{ P }{ S }{ Q }
}
%
\end{mathpar}

@@

\sectskip
\section{Conclusion}
\label{sec:concl}
\aftersectskip

@@

@@

\begin{figure*}
\setlength{\progwidth}{.24\linewidth}
  \centering

@@

\begin{figure*}
 \setlength{\progwidth}{.24\linewidth}
  \centering
  \begin{minipage}[b]{\progwidth}
    \begin{center}
   \begin{lstlisting}
  while (n>x) {
    if (m>y)
      y = y+1;
    else
      x = x+1;
  }
   \end{lstlisting}

@@

\begin{figure*}
 \setlength{\progwidth}{.24\linewidth}
  \centering
  \begin{minipage}[b]{\progwidth}
    \begin{center}
   \begin{lstlisting}
  x=0;
  while (x<n) {
    if (*) break;
    x=x+1;
  }
  while (x<n)
    x=x+1;
   \end{lstlisting}

@@

\begin{figure*}
 \setlength{\progwidth}{.24\linewidth}
  \centering
%
%
  \begin{minipage}[b]{\progwidth}
    \begin{center}
   \begin{lstlisting}
  assert n>0;
  assert m>0;
  va = n; vb = 0;
  while (va>0 && *) {
    if (vb<m) {
      vb=vb+1;
      va=va-1;
    } else {
      vb=vb-1;
      vb=0;
    }
  }
   \end{lstlisting}
$1 + 2|[0, n]|$
\\[.7\baselineskip]
      {\bf fig4\_2}
    \end{center}
  \end{minipage}
%
%
%
%
  \begin{minipage}[b]{\progwidth}
    \begin{center}
   \begin{lstlisting}
  assert (0<m);
  i = n;
  while (i>0 && *) {
    if (i<m)
      i=i-1;
    else
      i=i-m;
  }
   \end{lstlisting}
$|[0, n]|$
\\[.7\baselineskip]
      {\bf fig4\_4}
    \end{center}
  \end{minipage}
%
%
  \begin{minipage}[b]{\progwidth}
    \begin{center}
   \begin{lstlisting}
  assert (0 < m < n);
  i=m;
  while (0<i<n) {
    if (dir==fwd) i++;
    else i--;
  }
   \end{lstlisting}
$---$
\\[.7\baselineskip]
      {\bf fig4\_5}
    \end{center}
  \end{minipage}

@@

\begin{figure*}
 \setlength{\progwidth}{.28\linewidth}
  \centering
%
%
  \begin{minipage}[b]{\progwidth}
    \begin{center}
   \begin{lstlisting}
void count_down (int x) {
  int a = x;
  if (a>0) {
    a = a-1;
    count_down(a);
  }
}
int copy (int x, int y) {
  if (x>0) {
    x = x-1;
    y = y+1;
    y=copy(x,y);
  };
  return y;
}
void main (int x,int y) {
  y = copy (x,y);
  count_down(y);
}
   \end{lstlisting}
